### 2024-01-17 outputs/paper/paper.qmd
Update LLM and paper
Add a ChatGPT conversation about an error question to the LLM file

Create a table for COVID-19 case data in the paper file

---
title: "Exploring the Demographic Aspects of COVID-19 in Toronto: Unravelling Age and Gender Patterns"
subtitle: "Are the male elderly more susceptible to COVID-19?"
author: 
  - Amie Liu
thanks: "Code and data are available at: https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages("tidyverse")
# install.packages("here")
# install.packages("knitr")

library(tidyverse)
library(here)
library(knitr)

# Set the working directory to the project's root
setwd(here())

### Explore ###
# Read cleaned case data
# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore-1

cleaned_case_data <- 
  read_csv(
    "outputs/data/cleaned_case_data.csv", 
    show_col_types = FALSE)

# Read analyzed case data
# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore-1

analyzed_case_data <-
  read_csv(
    "outputs/data/analyzed_case_data.csv",
    show_col_types = FALSE)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

## COVID-19 Cases Data

```{r}
#| label: tbl-idcasetable
#| fig-cap: First six rows of COVID-19 cases by age groups and gender
#| echo: false
#| message: false
#| warning: false

# Reference: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
head(cleaned_case_data) |>
  kable(
    col.names = c("Case ID", "Age group", "Gender"),
    booktabs = TRUE
  )
```

## Number of Covid-19 Cases by Gender

## Number of Covid-19 Cases by Age Groups

## Distribution of Covid-19 Cases Across Genders and Age Groups

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-01-24 outputs/paper/paper.qmd
Update paper, references, llm usage, README
Complete data analysis, adding graphs and paragraphs to the paper file

Correct typo in the simulate, download, and clean data file

Add secondary research to the literature file

Add sketches to the sketches file

@@
 ---
-title: "Exploring the Demographic Aspects of COVID-19 in Toronto: Unravelling Age and Gender Patterns"
-subtitle: "Are the male elderly more susceptible to COVID-19?"
+title: "Exploring the Demographic Aspects of COVID-19 Cases in Toronto: Are the elderly males more susceptible to COVID-19?"
 author: 
   - Amie Liu
-thanks: "Code and data are available at: https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
+thanks: "Code and data are available at:https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Examining the age and gender patterns of COVID-19 cases in Toronto provides insights into the susceptibility of different demographic groups to the virus.
+This paper utilizes the dataset from Open Data Toronto, encompassing the period from January 22, 2020, to January 4, 2024.
+Contrary to the hypothesis that elderly males are more susceptible, the results reveal a pattern, with females aged 20-29 showing the highest susceptibility, exceeding 40,000 cases.
+This highlights the importance of understanding gender and age-specific patterns for effective public health interventions."
 format: pdf
@@
 # Read cleaned case data
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore-1
-
 cleaned_case_data <- 
@@
 # Read analyzed case data
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore-1
-
 analyzed_case_data <-
@@
 
-You can and should cross-reference sections and sub-sections.
+Global disease pandemics pose significant threats to public health and impact daily life worldwide.
+In December 2019, the outbreak of Coronavirus Disease 2019 (COVID-19) was identified and is caused by the severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) [@newDisease].
+This disease commonly manifests with respiratory symptoms similar to those of pneumonia, colds, or influenza, and some COVID-19 patients may experience severe illness [@aboutCovid].
+Researchers have found that COVID-19 cases tend to be more severe in the elderly, with males experiencing a higher severity rate than females, highlighting the influence of gender and age differences on the severity of COVID-19 [@genderDifferences].
+However, limited information is available regarding these gender and age differences that may influence susceptibility to COVID-19 [@pandemicEvolution].
+This leads to a question: “Are the elderly males more susceptible to COVID-19?”
 
-The remainder of this paper is structured as follows. @sec-data....
+Researchers have collaborated to provide a better understanding of infectious agents and develop techniques to mitigate outbreaks.
+Public health and social measures (PHSMs) have been a significant technique in decreasing the transmission of viruses with pandemic potential [@phsm].
+Understanding the demographic aspects of infection is crucial for exploring how different groups are affected and facilitating the formulation of effective Public Health and Social Measures (PHSMs).
+Moreover, this understanding can help with the resource allocation for vaccines and predict the demand for medical care across diverse demographic groups.
+
+Given the observation by public health officials that Toronto has a high level of COVID-19 activity [@influenzaActivity], I conducted an analysis of the relationship between COVID-19 cases and age groups, as well as genders, in Toronto.
+My results indicate a pattern of susceptibility in the distribution of COVID-19 cases in Toronto, where the majority of age groups show a higher number of female cases compared to males.
+Specifically, the group of females aged 20-29 has the highest number of COVID-19 cases—more than 40,000 cases.
+Contrary to the hypothesis, my data does not show that elderly males are more susceptible to COVID-19.
+The observed pattern may be influenced by factors beyond age and gender demographics.
+
+This paper is divided into the Data and Limitations sections.
+@sec-data Data delves into the various sources used for data collection, cleaning, and analyzing processes, leveraging a comprehensive dataset provided by Toronto Public Health on the City of Toronto's Open Data Portal [@covidCases].
+In addition, it provides a detailed analysis of the findings from the data.
+Moving forward, @sec-limitations Limitations reveal potential constraints, including considerations related to occupational exposure, geographical focus, variable measurement, and data accuracy, that might influence the interpretation of the main findings.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The data for this paper was obtained from the City of Toronto's Open Data Portal through the 'opendatatoronto' library [@openDataToronto].
+The details of the data cleaning process for the dataset 'COVID-19 Cases in Toronto' [@covidCases] are provided in @sec-appendix Appendix.
+The data analysis involved using the R programming language [@citeR] along with the 'tidyverse' [@citeTidyverse] 'here' [@citeHere], 'janitor' [@citeJanitor], 'knitr' [@citeKnitr] packages for data collection, cleaning, and analysis.
+Within the 'tidyverse' package [@citeTidyverse], I used 'readr' [@citeReadr], 'dplyr' [@citeDplyr], 'tidyr' [@citeTidyr], and 'ggplot2' [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating bar charts.
 
-## COVID-19 Cases Data
+## Covid-19 Cases by Gender
+
+Based on the bar chart of COVID-19 cases in Toronto by gender (see @fig-casebygender), the data reveals that females constitute the largest group, with 219,671 cases, comprising 53.4% of the total cases.
+Following closely, males represent the second-highest group, with 188,197 cases, accounting for 45.7% of the total cases.
+Additionally, 3,832 cases fall into the "Other" category, including transgender and not listed gender, making up 0.9% of the total cases.
+
+The gender breakdown of COVID-19 cases highlights a distribution pattern, underscoring the importance of integrating gender-specific considerations into public health interventions.
+There are no significant differences between female and male cases.
+However, the higher occurrence of cases among females suggests a need for targeted measurements that consider gender-specific factors.
+These findings may prompt health officials to communicate and implement outreach strategies, ensuring the effectiveness of preventive measures and vaccination campaigns across diverse gender groups.
+
+```{r}
+#| label: fig-casebygender
+#| fig-cap: Number of covid-19 cases by gender
+#| echo: false
+#| warning: false
+#| message: false
+
+# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
+analyzed_case_data |> 
+  ggplot(aes(x = gender)) +
+  geom_bar() +
+  geom_text(
+    stat = 'count',
+    aes(label = scales::percent(..count.. / sum(..count..)),
+        y = ..count..),
+    vjust = -0.5,
+    position = position_stack(vjust = 0.5)
+  ) +
+  theme_minimal() +
+  labs(x = "Gender",
+       y = "Number of covid-19 cases")
+```
+
+## Covid-19 Cases by Age Groups
+
+In @fig-casebyagegroup, the bar chart provides a clear depiction of the distribution of COVID-19 cases across various age groups.
+Particularly, the age group 20-29 has the highest number of cases, totaling 77,523 instances and representing 18.83% of the overall cases.
+This is closely followed by the age groups 30-39 and 0-19, with 73,400 (17.83%) and 55,142 (13.39%) cases, respectively.
+In the middle-age categories, 40-49 and 50-59 years old, there are 13.86% and 12.81% of the total cases, respectively.
+A gradual decline in COVID-19 cases is observed as the age increases, with the age group 60-69 at 8.62%, 70-79 at 5.73%, 80-89 at 5.51%, and the 90+ category at 3.40%.
+
+The concentration of cases in the age group 20-29 highlights the need for targeted outreach strategies for these younger people.
+Allocating resources strategically in areas with higher numbers of younger people may be an effective mitigation strategy.
+Furthermore, the comparatively lower number of cases in older age groups, particularly those aged 60 and above, suggests that these groups may have adopted and implemented effective preventive measures.
+
+```{r}
+#| label: fig-casebyagegroup
+#| fig-cap: Number of covid-19 cases by age groups
+#| echo: false
+#| warning: false
+#| message: false
+
+# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
+analyzed_case_data |> 
+  ggplot(aes(x = age_group)) +
+  geom_bar() +
+  geom_text(
+    stat = 'count',
+    aes(label = scales::percent(..count.. / sum(..count..)),
+        y = ..count..),
+    vjust = -0.5,
+    position = position_stack(vjust = 0.5)
+  ) +
+  theme_minimal() +
+  labs(x = "Age groups",
+       y = "Number of covid-19 cases")
+```
+
+## Distribution of Covid-19 Cases Across Genders and Age Groups
+
+As depicted in @fig-distribution, the majority of age groups show a higher number of female cases compared to male cases.
+However, exceptions to this trend are observed in the 0-19 and 70-79 age groups, where male cases exceed females.
+Moreover, the "Other" category consistently represents the lowest number of cases among all age groups.
+
+The data reveals that the highest incidence of COVID-19 cases is among females aged 20-29, with over 40,000 cases.
+The second-highest occurs in females aged 30-39, while the third-highest is observed among males in the 20-29 age group.
+These findings emphasize the significance of understanding age and gender-specific patterns.
+This paper seeks to address the question of whether elderly males are more susceptible to COVID-19; however, the data does not support this hypothesis, indicating that patterns of susceptibility may include influences beyond age and gender demographics.
+
+```{r}
+#| label: fig-distribution
+#| fig-cap: Distribution of covid-19 cases across genders and age groups
+#| echo: false
+#| warning: false
+#| message: false
+
+# Reference: https://rstudio-pubs-static.s3.amazonaws.com/215516_c2cee6b2885d419e9248e7eb7c87f887.html
+analyzed_case_data |> 
+  ggplot(aes(x = age_group, fill = gender)) +
+  geom_bar(position = "dodge") +
+  scale_x_discrete("Age groups") +
+  scale_fill_discrete(name = "Gender",
+                      breaks = c("Male", "Female", "Other")) +
+  theme_minimal() +
+  labs(y = "Number of covid-19 cases")
+```
+
+# Limitations {#sec-limitations}
+
+While this paper provides valuable insights into the age and gender patterns of COVID-19 cases in Toronto, several limitations should be taken into consideration, including occupational exposure, geographical focus, variable measurement, and data accuracy.
+
+## Occupational Exposure
+
+Occupational exposure is a key factor that may influence the demographic distribution of COVID-19 cases.
+Jobs involving close interaction with the public, such as those in healthcare and social assistance, could pose elevated risks of exposure.
+It is important to note that within these sectors, there is a higher prevalence of female workers compared to male workers and those aged 25-54 compared to other age groups [@healthCare].
+
+## Geographical focus
+
+The focus on Toronto in this paper may limit the generalizability of research findings to other areas or populations, given that diverse areas may exhibit distinct patterns.
+Therefore, the results presented here may not be universally applicable.
+The scope of this paper is specifically targeted at Toronto, and other areas are beyond the intended scope of this paper.
+
+## Variable Measurement
+
+The comparison of COVID-19 cases by gender and age group in Toronto raises considerations about the representativeness of the observed patterns with the population.
+The resident population data indicates variations in the proportions of males and females, with females slightly outnumbering males [@residentPopulation], and the age distribution, particularly the largest population (20-29 age group) [@populationOfToronto], should be noticed.
+This difference in the underlying population composition may have implications for the interpretation of COVID-19 cases.
+Although the population data aligns with the observed COVID-19 case data, it is important to know that the general population's proportions of gender and age groups should be taken into account.
+This limitation highlights the importance of considering demographic factors, such as comparing the resident population to the COVID-19 case data, to ensure a more comprehensive pattern of my findings.
+
+## Data Accuracy
+
+The data used for this analysis may not fully capture all COVID-19 cases within the population of Toronto.
+The City of Toronto highlights a significant change in testing guidance by the Ministry of Health starting in 2021 [@covidDashboards], leading to a limitation in the available data, as only specific subsets of the population are being tested.
+Certain data may no longer accurately represent the broader population [@covidDashboards].
+Moreover, certain people infected with COVID-19 may not report to the Ministry of Health, and those who show no symptoms could be overlooked in the dataset.
+This limitation leads to the challenge of accurately assessing the data on COVID-19 cases.
+
+Future research could address these limitations by exploring other demographic variables and considering a wider geographical scope.
+This would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases in diverse populations.
+
+\newpage
+
+# Appendix {#sec-appendix}
+
+## Data Cleaning
+
+The dataset 'COVID-19 Cases in Toronto'  [@covidCases] was utilized and obtained from 'opendatatoronto' library [@openDataToronto].
+It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more within the COVID-19 cases in Toronto.
+The data cover the period from January 22, 2020, to January 4, 2024, and are structured in CSV file format.
+
+I conducted a data-cleaning procedure, focusing on variable selection, simplifying variable names, and handling missing values.
+The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender.
+If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable).
+The total number of rows has been reduced from 412,122 to 411,700.
 
@@
 #| label: tbl-idcasetable
-#| fig-cap: First six rows of COVID-19 cases by age groups and gender
+#| tbl-cap: First six rows of COVID-19 cases by age groups and gender
 #| echo: false
@@
 
-# Reference: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
 head(cleaned_case_data) |>
   kable(
-    col.names = c("Case ID", "Age group", "Gender"),
-    booktabs = TRUE
+    col.names = c("Case ID", "Age group", "Gender")
   )
-```
-
-## Number of Covid-19 Cases by Gender
-
-## Number of Covid-19 Cases by Age Groups
-
-## Distribution of Covid-19 Cases Across Genders and Age Groups
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
 ```

### 2024-01-24 scripts/00-simulate_data.R

@@ -16,7 +16,7 @@ library(janitor)
 #### Simulate data ####
 # Number of cases and types of gender and age groups for simulating data
 case_num <- 412122
-simulated_gender <- c("Male", "Female")
+simulated_gender <- c("Male", "Female", "Other")
 simulated_age <- c("0-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+")
 
 # Generate a dataset containing COVID-19 cases and genders
@@ -46,10 +46,10 @@ simulated_data
 
 #### Test data ####
 # Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate-2
-# Check if "gender" contains 2 genders
+# Check if "gender" contains 3 genders
 simulated_data$gender |>
   unique() |>
-  length() == 2
+  length() == 3
 
 # Check if "age_group" contains 9 groups
 simulated_data$age_group |>

### 2024-01-24 scripts/01-download_data.R

@@ -7,13 +7,11 @@
 # Pre-requisites: None
 
 #### Workspace setup ####
-#install.packages("opendatatoronto")
-#install.packages("dplyr")
-#install.packages("readr")
+# install.packages("opendatatoronto")
+# install.packages("tidyverse")
 
 library(opendatatoronto)
-library(dplyr)
-library(readr)
+library(tidyverse)
 
 #### Download COVID-19 cases in Toronto data ####
 # Code from: https://open.toronto.ca/dataset/covid-19-cases-in-toronto/

### 2024-01-24 scripts/02-data_cleaning.R

@@ -7,7 +7,8 @@
 # Pre-requisites: None
 
 #### Workspace setup ####
-#install.packages("tidyverse")
+# install.packages("tidyverse")
+# install.packages("janitor")
 
 library(tidyverse)
 library(janitor)
@@ -57,7 +58,7 @@ cleaned_case_data <-
          ) |>
   tidyr::drop_na()
 
-# Analysis data, including age group and gender
+# Analyzed data, including age group and gender
 analyzed_case_data <-
   cleaned_case_data |>
   select(age_group, gender)

### 2024-01-25 outputs/paper/paper.qmd
Update Paper
Add appendix
Update paper and llm
Add table of contents and discussion section in paper

Add ChatGPT usage in llm file

@@
 format: pdf
+toc: true
 number-sections: true
@@
 @sec-data Data delves into the various sources used for data collection, cleaning, and analyzing processes, leveraging a comprehensive dataset provided by Toronto Public Health on the City of Toronto's Open Data Portal [@covidCases].
-In addition, it provides a detailed analysis of the findings from the data.
-Moving forward, @sec-limitations Limitations reveal potential constraints, including considerations related to occupational exposure, geographical focus, variable measurement, and data accuracy, that might influence the interpretation of the main findings.
+In addition, @sec-results Results provides a detailed analysis of the findings from the data.
+Moving forward, @sec-discussion Discussion provides an understanding of the demographic patterns of COVID-19 cases, highlights key findings, and reveals potential limitations, including considerations related to occupational exposure, geographical focus, variable measurement, and data accuracy, that might influence the interpretation of the main findings.
 
@@
 The data for this paper was obtained from the City of Toronto's Open Data Portal through the 'opendatatoronto' library [@openDataToronto].
-The details of the data cleaning process for the dataset 'COVID-19 Cases in Toronto' [@covidCases] are provided in @sec-appendix Appendix.
 The data analysis involved using the R programming language [@citeR] along with the 'tidyverse' [@citeTidyverse] 'here' [@citeHere], 'janitor' [@citeJanitor], 'knitr' [@citeKnitr] packages for data collection, cleaning, and analysis.
@@
 
-## Covid-19 Cases by Gender
+## COVID-19 Cases Data
+
+The dataset 'COVID-19 Cases in Toronto'  [@covidCases] was utilized and obtained from 'opendatatoronto' library [@openDataToronto].
+It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more within the COVID-19 cases in Toronto.
+The data cover the period from January 22, 2020, to January 4, 2024, and are structured in CSV file format.
+
+The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender.
+If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable).
+The total number of rows has been reduced from 412,122 to 411,700.
+
+```{r}
+#| label: tbl-idcasetable
+#| tbl-cap: Sample of cleaned COVID-19 cases data
+#| echo: false
+#| message: false
+#| warning: false
+
+head(cleaned_case_data) |>
+  kable(
+    col.names = c("Case ID", "Age group", "Gender")
+  )
+```
+
+# Results {#sec-results}
+
+## COVID-19 Cases by Gender
 
@@
 
-## Covid-19 Cases by Age Groups
+## COVID-19 Cases by Age Groups
 
@@
 
-## Distribution of Covid-19 Cases Across Genders and Age Groups
+## Distribution of COVID-19 Cases Across Genders and Age Groups
 
@@
 
-# Limitations {#sec-limitations}
+# Discussion {#sec-discussion}
+
+## Interpretation of Findings
+The examination of COVID-19 cases by gender reveals that females represent 53.4% of the cases in Toronto and males represent 45.7% (see @fig-casebygender), which does not show significant differences between female and male cases.
+The age-stratified analysis of COVID-19 cases shows that the 20-29 age group is most affected by COVID-19, accounting for 18.83% of the overall cases (see @fig-casebyagegroup).
+Conversely, older age groups (40+) have lower case numbers.
+Moreover, the intersectional analysis of gender and age groups shows that most age groups have a higher number of female cases, with exceptions in the 0-19 and 70-79 age groups where male cases are higher than female cases (see @fig-distribution).
+These findings do not support the hypothetical question: "Are elderly males more susceptible to COVID-19?" Additional demographic variables, such as socioeconomic status and cultural factors, may influence COVID-19 patterns in diverse populations.
+
+## Limitations
 
@@
 
-## Occupational Exposure
+### Occupational Exposure
 
@@
 
-## Geographical focus
+### Geographical focus
 
@@
 
-## Variable Measurement
+### Variable Measurement
 
@@
 
-## Data Accuracy
+### Data Accuracy
 
@@
 
+## Next Steps
+
 Future research could address these limitations by exploring other demographic variables and considering a wider geographical scope.
-This would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases in diverse populations.
-
-\newpage
-
-# Appendix {#sec-appendix}
-
-## Data Cleaning
-
-The dataset 'COVID-19 Cases in Toronto'  [@covidCases] was utilized and obtained from 'opendatatoronto' library [@openDataToronto].
-It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more within the COVID-19 cases in Toronto.
-The data cover the period from January 22, 2020, to January 4, 2024, and are structured in CSV file format.
-
-I conducted a data-cleaning procedure, focusing on variable selection, simplifying variable names, and handling missing values.
-The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender.
-If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable).
-The total number of rows has been reduced from 412,122 to 411,700.
-
-```{r}
-#| label: tbl-idcasetable
-#| tbl-cap: First six rows of COVID-19 cases by age groups and gender
-#| echo: false
-#| message: false
-#| warning: false
-
-head(cleaned_case_data) |>
-  kable(
-    col.names = c("Case ID", "Age group", "Gender")
-  )
-```
+This exploration would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases across diverse populations.
+Future research endeavors have the potential to refine public health strategies, tailor interventions to specific population needs, and contribute substantially to the global understanding of the multifaceted nature of the COVID-19 pandemic.
 

### 2024-01-26 outputs/paper/paper.qmd
Update paper, simulate data, download data, data cleaning, README, llm usage, and add test data based on the peer review suggestions
Fix typos, add pre-requisites, additional comments to the code, and add details to the findings

Add tests for the cleaned data

@@
 The dataset 'COVID-19 Cases in Toronto'  [@covidCases] was utilized and obtained from 'opendatatoronto' library [@openDataToronto].
-It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more within the COVID-19 cases in Toronto.
+It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more.
 The data cover the period from January 22, 2020, to January 4, 2024, and are structured in CSV file format.
@@
 The primary variables selected for analysis were chosen based on their relevance to my research question, relating to age and gender.
-If one categorical value was missing, then the entire row was removed Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable).
+If one categorical value was missing, then the entire row was removed.
+Therefore, the cleaned COVID-19 cases dataset now includes three variables: case ID, gender, and age group (see @tbl-idcasetable).
 The total number of rows has been reduced from 412,122 to 411,700.
@@
 
+# Display the first six rows of the "cleaned_case_data" dataset
 head(cleaned_case_data) |>
   kable(
+    # Give each column a name
     col.names = c("Case ID", "Age group", "Gender")
@@
 
+# Create a bar chart to visualize the distribution of COVID-19 cases by gender based on the analyzed_case_data dataset
 # Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
@@
   ggplot(aes(x = gender)) +
-  geom_bar() +
+  # Change bar color to green
+  geom_bar(fill = "#D4E7C5") +
   geom_text(
+    # Use the 'count' statistic to calculate counts for each category
     stat = 'count',
+    # Use percentage labels based on count and set y position of text labels
     aes(label = scales::percent(..count.. / sum(..count..)),
         y = ..count..),
+    # Adjust vertical position of text labels
     vjust = -0.5,
+    # Stack text labels on top of bars
     position = position_stack(vjust = 0.5)
   ) +
+  # Apply a minimal theme to the chart
   theme_minimal() +
+  # Label x-axis as "Gender" and y-axis as "Number of COVID-19 cases"
   labs(x = "Gender",
-       y = "Number of covid-19 cases")
+       y = "Number of COVID-19 cases")
 ```
 
-## COVID-19 Cases by Age Groups
+## COVID-19 Cases by Age Group
 
@@
 #| label: fig-casebyagegroup
-#| fig-cap: Number of covid-19 cases by age groups
+#| fig-cap: Number of covid-19 cases by age group
 #| echo: false
@@
 
+# Create a bar chart to visualize the distribution of COVID-19 cases by age group based on the analyzed_case_data dataset
 # Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
@@
   ggplot(aes(x = age_group)) +
-  geom_bar() +
+  # Change bar color to green
+  geom_bar(fill = "#D4E7C5") +
   geom_text(
+    # Use the 'count' statistic to calculate counts for each category
     stat = 'count',
+    # Use percentage labels based on count and set y position of text labels
     aes(label = scales::percent(..count.. / sum(..count..)),
         y = ..count..),
+    # Adjust vertical position of text labels
     vjust = -0.5,
-    position = position_stack(vjust = 0.5)
+    # Stack text labels on top of bars
+    position = position_stack(vjust = 0.5),
+    # Change text size to 3
+    size = 3
   ) +
+  # Apply a minimal theme to the chart
   theme_minimal() +
-  labs(x = "Age groups",
-       y = "Number of covid-19 cases")
+  # Label x-axis as "Age group" and y-axis as "Number of COVID-19 cases"
+  labs(x = "Age group",
+       y = "Number of COVID-19 cases")
 ```
@@
 analyzed_case_data |> 
+  # Create a ggplot object with age_group on the x-axis and gender as fill color
   ggplot(aes(x = age_group, fill = gender)) +
+  # Add a grouped bar layer to create the grouped bar chart
   geom_bar(position = "dodge") +
-  scale_x_discrete("Age groups") +
+  # Scale the x-axis labels to display "Age group"
+  scale_x_discrete("Age group") +
+  # Scale the fill color to display "Gender" and specify the break points: "Male", "Female", "Other"
   scale_fill_discrete(name = "Gender",
                       breaks = c("Male", "Female", "Other")) +
+  # Apply a minimal theme to the chart
   theme_minimal() +
-  labs(y = "Number of covid-19 cases")
+  # Label y-axis as "Number of COVID-19 cases"
+  labs(y = "Number of COVID-19 cases")
 ```
@@
 The examination of COVID-19 cases by gender reveals that females represent 53.4% of the cases in Toronto and males represent 45.7% (see @fig-casebygender), which does not show significant differences between female and male cases.
+This indicates that both genders are equally affected by COVID-19 in Toronto.
+Although the overall proportions may appear similar, there may be differences in the severity of the cases with each gender that need further investigation.
+
 The age-stratified analysis of COVID-19 cases shows that the 20-29 age group is most affected by COVID-19, accounting for 18.83% of the overall cases (see @fig-casebyagegroup).
-Conversely, older age groups (40+) have lower case numbers.
+Conversely, older age groups, particularly those aged 40 and above, have lower case numbers.
+This suggests a potential age-related pattern in COVID-19 susceptibility or exposure and highlights the importance of targeted age-specific interventions and public health measures to effectively mitigate the virus's spread.
+
 Moreover, the intersectional analysis of gender and age groups shows that most age groups have a higher number of female cases, with exceptions in the 0-19 and 70-79 age groups where male cases are higher than female cases (see @fig-distribution).
-These findings do not support the hypothetical question: "Are elderly males more susceptible to COVID-19?" Additional demographic variables, such as socioeconomic status and cultural factors, may influence COVID-19 patterns in diverse populations.
+This highlights the complexity of COVID-19 epidemiology.
+Further research is needed to explore the underlying factors that cause this pattern.
+
+These findings do not support the hypothetical question: "Are elderly males more susceptible to COVID-19?" Instead, the data suggests that COVID-19 patterns are influenced by factors beyond age and gender demographics.
+Additional demographic variables, such as socioeconomic status and cultural factors, may influence COVID-19 patterns in diverse populations.
 
@@
 Future research could address these limitations by exploring other demographic variables and considering a wider geographical scope.
-This exploration would provide a more comprehensive understanding of the factors that influence the patterns of COVID-19 cases across diverse populations.
-Future research endeavors have the potential to refine public health strategies, tailor interventions to specific population needs, and contribute substantially to the global understanding of the multifaceted nature of the COVID-19 pandemic.
+This exploration would provide a more detailed understanding of the factors that influence the patterns of COVID-19 cases across diverse populations.
+This could have the potential to refine public health strategies, inform targeted public health interventions, and contribute substantially to the global understanding of the COVID-19 pandemic.
 

### 2024-01-26 scripts/00-simulate_data.R

@@ -51,6 +51,9 @@ simulated_data$gender |>
   unique() |>
   length() == 3
 
+# Check if "gender" contains 3 specified variables: "Male", "Female", and "Other"
+all(unique(simulated_data$gender) %in% c("Male", "Female", "Other"))
+
 # Check if "age_group" contains 9 groups
 simulated_data$age_group |>
   unique() |>

### 2024-01-26 scripts/01-download_data.R

@@ -14,7 +14,7 @@ library(opendatatoronto)
 library(tidyverse)
 
 #### Download COVID-19 cases in Toronto data ####
-# Code from: https://open.toronto.ca/dataset/covid-19-cases-in-toronto/
+# COVID-19 cases Dataset and code are from: https://open.toronto.ca/dataset/covid-19-cases-in-toronto/
 # Get all resources from this package
 resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
 

### 2024-01-26 scripts/02-data_cleaning.R

@@ -4,7 +4,9 @@
 # Date: 16 January 2024
 # Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites:
+#   00-simulate_data.R
+#   01-download_data.R
 
 #### Workspace setup ####
 # install.packages("tidyverse")
@@ -21,7 +23,9 @@ raw_case_data <- read_csv("inputs/data/raw_case_data.csv")
 # Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 cleaned_case_data <-
   raw_case_data |>
+  # Clean the column names
   clean_names() |>
+  # Transform the existing "age_group" values into standardized age groups
   mutate(
     age_group =
       case_match(
@@ -37,6 +41,7 @@ cleaned_case_data <-
         "90 and older" ~ "90+",
       )
   ) |>
+  # Transform the existing "gender" values into standardized categories ("Female", "Male", or "Other")
   mutate(
     client_gender =
       case_match(
@@ -52,10 +57,12 @@ cleaned_case_data <-
         "NOT LISTED, PLEASE SPECIFY" ~ "Other",
       )
   ) |>
+  # Rename the columns
   select(assigned_id, age_group, client_gender) |>
   rename(case_id = assigned_id,
          gender = client_gender
          ) |>
+  # Remove any rows with missing values (NA) in the dataset
   tidyr::drop_na()
 
 # Analyzed data, including age group and gender

### 2024-01-26 scripts/03-test_data.R

@@ -1,0 +1,54 @@
+#### Preamble ####
+# Purpose: Test cleaned COVID-19 cases data
+# Author: Amie Liu
+# Date: 25 January 2024
+# Contact: amie.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+#   00-simulate_data.R
+#   01-download_data.R
+#   02-data_cleaning.R
+
+#### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("janitor")
+
+library(tidyverse)
+library(janitor)
+
+#### Test data ####
+# Read data
+cleaned_case_data <- read_csv("outputs/data/cleaned_case_data.csv")
+
+# Test cleaned data
+# Check if all column names match the specified names and order: "case_id", "age_group", "gender"
+all(colnames(cleaned_case_data) == c("case_id", "age_group", "gender"))
+
+# Check if the minimum value of "case_id" is 1
+cleaned_case_data$case_id |> min() == 1
+
+# Check if the length of "case_id" is 411700
+cleaned_case_data$case_id |> length() == 411700
+
+# Check if the class of "case_id" is numeric
+cleaned_case_data$case_id |> class() == "numeric"
+
+# Check if "age_group" contains 9 groups
+cleaned_case_data$age_group |>
+  unique() |>
+  length() == 9
+
+# Check if the length of "age_group" is 411700
+cleaned_case_data$age_group |> length() == 411700
+
+# Check if "gender" contains 3 genders
+cleaned_case_data$gender |>
+  unique() |>
+  length() == 3
+
+# Check if "gender" contains 3 specified variables: "Male", "Female", and "Other"
+all(unique(cleaned_case_data$gender) %in% c("Male", "Female", "Other"))
+
+# Check if the length of "gender" is 411700
+cleaned_case_data$gender |> length() == 411700
+

### 2025-06-17 scripts/00-simulate_data.R
changed scripts -- cleaned_case_data changed?

@@ -1,66 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates the dataset of COVID-19 cases in Toronto
-# Author: Amie Liu
-# Date: 16 January 2024
-# Contact: amie.liu@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-# install.packages("tidyverse")
-# install.packages("janitor")
-
-library(tidyverse)
-library(janitor)
-
-#### Simulate data ####
-# Number of cases and types of gender and age groups for simulating data
-case_num <- 412122
-simulated_gender <- c("Male", "Female", "Other")
-simulated_age <- c("0-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+")
-
-# Generate a dataset containing COVID-19 cases and genders
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate
-# Set a random seed
-set.seed(123)
-
-simulated_data <-
-  tibble(
-    # Use 1 through case_num to represent each case_id
-    case_id = c(1:case_num),
-    # Randomly pick an option for gender, with replacement, case_num times
-    gender = sample(
-      x = simulated_gender,
-      size = case_num,
-      replace = TRUE
-    ),
-    # Randomly pick an option for age group, with replacement, case_num times
-    age_group = sample(
-      x = simulated_age,
-      size = case_num,
-      replace = TRUE
-    )
-  )
-
-simulated_data
-
-#### Test data ####
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate-2
-# Check if "gender" contains 3 genders
-simulated_data$gender |>
-  unique() |>
-  length() == 3
-
-# Check if "gender" contains 3 specified variables: "Male", "Female", and "Other"
-all(unique(simulated_data$gender) %in% c("Male", "Female", "Other"))
-
-# Check if "age_group" contains 9 groups
-simulated_data$age_group |>
-  unique() |>
-  length() == 9
-
-# Check if the minimum value of "case_id" is 1
-simulated_data$case_id |> min() == 1
-
-# Check if the length of "case_id" is 412122
-simulated_data$case_id |> length() == 412122

### 2025-06-17 scripts/00-simulate_data.py

@@ -1,0 +1,45 @@
+#### Preamble ####
+# Purpose: Simulates the dataset of COVID-19 cases in Toronto
+# Author: Amie Liu
+# Date: 16 January 2024
+# Contact: amie.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+import numpy as np
+import pandas as pd
+
+#### Simulate data ####
+# Number of cases and types of gender and age groups for simulating data
+case_num = 412122
+simulated_gender = ["Male", "Female", "Other"]
+simulated_age = ["0-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"]
+
+# Set a random seed
+np.random.seed(123)
+
+# Generate simulated data
+simulated_data = pd.DataFrame({
+    "case_id": np.arange(1, case_num + 1),
+    "gender": np.random.choice(simulated_gender, size=case_num, replace=True),
+    "age_group": np.random.choice(simulated_age, size=case_num, replace=True)
+})
+
+print(simulated_data)
+
+#### Test data ####
+# Check if "gender" contains 3 genders
+print(len(simulated_data["gender"].unique()) == 3)
+
+# Check if "gender" contains 3 specified variables: "Male", "Female", and "Other"
+print(set(simulated_data["gender"].unique()).issubset(set(["Male", "Female", "Other"])))
+
+# Check if "age_group" contains 9 groups
+print(len(simulated_data["age_group"].unique()) == 9)
+
+# Check if the minimum value of "case_id" is 1
+print(simulated_data["case_id"].min() == 1)
+
+# Check if the length of "case_id" is 412122
+print(len(simulated_data["case_id"]) == 412122)

### 2025-06-17 scripts/01-download_data.R

@@ -1,33 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from the Open Data Toronto portal
-# Author: Amie Liu
-# Date: 16 January 2024
-# Contact: amie.liu@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: None
-
-#### Workspace setup ####
-# install.packages("opendatatoronto")
-# install.packages("tidyverse")
-
-library(opendatatoronto)
-library(tidyverse)
-
-#### Download COVID-19 cases in Toronto data ####
-# COVID-19 cases Dataset and code are from: https://open.toronto.ca/dataset/covid-19-cases-in-toronto/
-# Get all resources from this package
-resources <- list_package_resources("64b54586-6180-4485-83eb-81e8fae3b8fe")
-
-# Identify datastore resources in csv format
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# Load the COVID-19 cases.csv
-raw_case_data <- filter(datastore_resources, row_number()==2) |> get_resource()
-
-raw_case_data
-
-#### Save data ####
-write_csv(raw_case_data, "inputs/data/raw_data.csv")
-
-# Rename the file for raw data
-file.rename("inputs/data/raw_data.csv", "inputs/data/raw_case_data.csv")

### 2025-06-17 scripts/01-download_data.py

@@ -1,0 +1,42 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from the Open Data Toronto portal
+# Author: Amie Liu
+# Date: 16 January 2024
+# Contact: amie.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: None
+
+#### Workspace setup ####
+import pandas as pd
+from io import StringIO
+from datetime import datetime
+
+#### Download COVID-19 cases in Toronto data ####
+import requests
+
+# Toronto Open Data is stored in a CKAN instance. It's APIs are documented here:
+# https://docs.ckan.org/en/latest/api/
+
+# To hit our API, you'll be making requests to:
+base_url = "https://ckan0.cf.opendata.inter.prod-toronto.ca"
+
+# Datasets are called "packages". Each package can contain many "resources"
+# To retrieve the metadata for this package and its resources, use the package name in this page's URL:
+url = base_url + "/api/3/action/package_show"
+params = { "id": "covid-19-cases-in-toronto"}
+package = requests.get(url, params = params).json()
+
+# To get resource data:
+for idx, resource in enumerate(package["result"]["resources"]):
+
+       # for datastore_active resources:
+       if resource["datastore_active"]:
+            # To get all records in CSV format:
+            url = base_url + "/datastore/dump/" + resource["id"]
+            resource_dump_data = requests.get(url)
+
+            # Writing raw data to CSV
+            raw_case_data = pd.read_csv(StringIO(resource_dump_data.text))
+            raw_case_data.to_csv("inputs/data/raw_case_data.csv")
+            
+

### 2025-06-17 scripts/02-data_cleaning.R

@@ -1,84 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw COVID-19 cases data
-# Author: Amie Liu
-# Date: 16 January 2024
-# Contact: amie.liu@mail.utoronto.ca
-# License: MIT
-# Pre-requisites:
-#   00-simulate_data.R
-#   01-download_data.R
-
-#### Workspace setup ####
-# install.packages("tidyverse")
-# install.packages("janitor")
-
-library(tidyverse)
-library(janitor)
-
-#### Clean data ####
-# Read data
-raw_case_data <- read_csv("inputs/data/raw_case_data.csv")
-
-# Clean data, including case ID, age group, and gender
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-cleaned_case_data <-
-  raw_case_data |>
-  # Clean the column names
-  clean_names() |>
-  # Transform the existing "age_group" values into standardized age groups
-  mutate(
-    age_group =
-      case_match(
-        age_group,
-        "19 and younger" ~ "0-19",
-        "20 to 29 Years" ~ "20-29",
-        "30 to 39 Years" ~ "30-39",
-        "40 to 49 Years" ~ "40-49",
-        "50 to 59 Years" ~ "50-59",
-        "60 to 69 Years" ~ "60-69",
-        "70 to 79 Years" ~ "70-79",
-        "80 to 89 Years" ~ "80-89",
-        "90 and older" ~ "90+",
-      )
-  ) |>
-  # Transform the existing "gender" values into standardized categories ("Female", "Male", or "Other")
-  mutate(
-    client_gender =
-      case_match(
-        client_gender,
-        "FEMALE" ~ "Female",
-        "MALE" ~ "Male",
-        "UNKNOWN" ~ "Other",
-        "NON-BINARY" ~ "Other",
-        "OTHER" ~ "Other",
-        "TRANSGENDER" ~ "Other",
-        "TRANS WOMAN" ~ "Other",
-        "TRANS MAN" ~ "Other",
-        "NOT LISTED, PLEASE SPECIFY" ~ "Other",
-      )
-  ) |>
-  # Rename the columns
-  select(assigned_id, age_group, client_gender) |>
-  rename(case_id = assigned_id,
-         gender = client_gender
-         ) |>
-  # Remove any rows with missing values (NA) in the dataset
-  tidyr::drop_na()
-
-# Analyzed data, including age group and gender
[…]

### 2025-06-17 scripts/02-data_cleaning.py

@@ -1,0 +1,74 @@
+#### Preamble ####
+# Purpose: Cleans the raw COVID-19 cases data
+# Author: Amie Liu
+# Date: 16 January 2024
+# Contact: amie.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+#   00-simulate_data.py
+#   01-download_data.py
+
+#### Workspace setup ####
+import pandas as pd
+import numpy as np
+
+#### Clean data ####
+# Read raw data
+raw_case_data = pd.read_csv("inputs/data/raw_case_data.csv")
+
+# Standardize column names to snake_case
+raw_case_data.columns = (
+    raw_case_data.columns
+    .str.strip()
+    .str.lower()
+    .str.replace(" ", "_")
+    .str.replace("-", "_")
+)
+
+# Map age group values
+age_group_mapping = {
+    "19 and younger": "0-19",
+    "20 to 29 Years": "20-29",
+    "30 to 39 Years": "30-39",
+    "40 to 49 Years": "40-49",
+    "50 to 59 Years": "50-59",
+    "60 to 69 Years": "60-69",
+    "70 to 79 Years": "70-79",
+    "80 to 89 Years": "80-89",
+    "90 and older": "90+"
+}
+
+# Map gender values
+gender_mapping = {
+    "FEMALE": "Female",
+    "MALE": "Male",
+    "UNKNOWN": "Other",
+    "NON-BINARY": "Other",
+    "OTHER": "Other",
+    "TRANSGENDER": "Other",
+    "TRANS WOMAN": "Other",
+    "TRANS MAN": "Other",
+    "NOT LISTED, PLEASE SPECIFY": "Other"
+}
+
+# Apply transformations
+cleaned_case_data = raw_case_data.copy()
+cleaned_case_data["age_group"] = cleaned_case_data["age_group"].map(age_group_mapping)
+cleaned_case_data["client_gender"] = cleaned_case_data["client_gender"].map(gender_mapping)
+
+# Select and rename columns
+cleaned_case_data = cleaned_case_data[["assigned_id", "age_group", "client_gender"]].rename(
+    columns={"assigned_id": "case_id", "client_gender": "gender"}
+)
+
+# Drop rows with any missing values
+cleaned_case_data = cleaned_case_data.dropna()
+
+# Select for analysis
+analyzed_case_data = cleaned_case_data[["age_group", "gender"]]
+
+# Save cleaned data
[…]

### 2025-06-17 scripts/03-test_data.R

@@ -1,54 +1,0 @@
-#### Preamble ####
-# Purpose: Test cleaned COVID-19 cases data
-# Author: Amie Liu
-# Date: 25 January 2024
-# Contact: amie.liu@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-#   00-simulate_data.R
-#   01-download_data.R
-#   02-data_cleaning.R
-
-#### Workspace setup ####
-# install.packages("tidyverse")
-# install.packages("janitor")
-
-library(tidyverse)
-library(janitor)
-
-#### Test data ####
-# Read data
-cleaned_case_data <- read_csv("outputs/data/cleaned_case_data.csv")
-
-# Test cleaned data
-# Check if all column names match the specified names and order: "case_id", "age_group", "gender"
-all(colnames(cleaned_case_data) == c("case_id", "age_group", "gender"))
-
-# Check if the minimum value of "case_id" is 1
-cleaned_case_data$case_id |> min() == 1
-
-# Check if the length of "case_id" is 411700
-cleaned_case_data$case_id |> length() == 411700
-
-# Check if the class of "case_id" is numeric
-cleaned_case_data$case_id |> class() == "numeric"
-
-# Check if "age_group" contains 9 groups
-cleaned_case_data$age_group |>
-  unique() |>
-  length() == 9
-
-# Check if the length of "age_group" is 411700
-cleaned_case_data$age_group |> length() == 411700
-
-# Check if "gender" contains 3 genders
-cleaned_case_data$gender |>
-  unique() |>
-  length() == 3
-
-# Check if "gender" contains 3 specified variables: "Male", "Female", and "Other"
-all(unique(cleaned_case_data$gender) %in% c("Male", "Female", "Other"))
-
-# Check if the length of "gender" is 411700
-cleaned_case_data$gender |> length() == 411700
-

### 2025-06-17 scripts/02-data_cleaning.py
fixed analyzed_data save

@@ -71,4 +71,4 @@ analyzed_case_data = cleaned_case_data[["age_group", "gender"]]
 cleaned_case_data.to_csv("outputs/data/cleaned_case_data.csv", index=False)
 
 # Save analyzed data
-analyzed_case_data.to_csv("analyzed_case_data.csv", index=False)
+analyzed_case_data.to_csv("outputs/data/analyzed_case_data.csv", index=False)

### 2025-06-17 scripts/03-test_data.py

@@ -5,50 +5,50 @@
 # Contact: amie.liu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-#   00-simulate_data.R
-#   01-download_data.R
-#   02-data_cleaning.R
+#   00-simulate_data.py
+#   01-download_data.py
+#   02-data_cleaning.py
 
 #### Workspace setup ####
-# install.packages("tidyverse")
-# install.packages("janitor")
+import unittest
+import pandas as pd
+import numpy as np
 
-library(tidyverse)
-library(janitor)
+class TestCleanedCovidData(unittest.TestCase):
 
-#### Test data ####
-# Read data
-cleaned_case_data <- read_csv("outputs/data/cleaned_case_data.csv")
+    @classmethod
+    def setUpClass(cls):
+        cls.df = pd.read_csv("outputs/data/cleaned_case_data.csv")
 
-# Test cleaned data
-# Check if all column names match the specified names and order: "case_id", "age_group", "gender"
-all(colnames(cleaned_case_data) == c("case_id", "age_group", "gender"))
+    def test_column_names_and_order(self):
+        expected_columns = ["case_id", "age_group", "gender"]
+        self.assertEqual(list(self.df.columns), expected_columns, "Column names or order are incorrect")
 
-# Check if the minimum value of "case_id" is 1
-cleaned_case_data$case_id |> min() == 1
+    def test_case_id_min_value(self):
+        self.assertEqual(self.df["case_id"].min(), 1, "Minimum case_id is not 1")
 
-# Check if the length of "case_id" is 411700
-cleaned_case_data$case_id |> length() == 411700
+    def test_case_id_length(self):
+        self.assertEqual(len(self.df["case_id"]), 411700, "case_id length is not 411700")
 
-# Check if the class of "case_id" is numeric
-cleaned_case_data$case_id |> class() == "numeric"
+    def test_case_id_is_numeric(self):
+        self.assertTrue(np.issubdtype(self.df["case_id"].dtype, np.number), "case_id is not numeric")
 
-# Check if "age_group" contains 9 groups
-cleaned_case_data$age_group |>
-  unique() |>
-  length() == 9
+    def test_age_group_unique_count(self):
+        self.assertEqual(self.df["age_group"].nunique(), 9, "age_group does not contain 9 unique values")
 
-# Check if the length of "age_group" is 411700
-cleaned_case_data$age_group |> length() == 411700
+    def test_age_group_length(self):
[…]

### 2025-06-17 outputs/paper/paper.qmd
changed scripts

@@
 
-```{r}
+```{python}
 #| include: false
@@
 
-# install.packages("tidyverse")
-# install.packages("here")
-# install.packages("knitr")
+#### Workspace setup ####
+import pandas as pd
 
-library(tidyverse)
-library(here)
-library(knitr)
+#### Explore ####
+# Read cleaned case data
+cleaned_case_data = pd.read_csv("../data/cleaned_case_data.csv")
+analyzed_case_data = pd.read_csv("../data/analyzed_case_data.csv")
 
-# Set the working directory to the project's root
-setwd(here())
-
-### Explore ###
-# Read cleaned case data
-cleaned_case_data <- 
-  read_csv(
-    "outputs/data/cleaned_case_data.csv", 
-    show_col_types = FALSE)
-
-# Read analyzed case data
-analyzed_case_data <-
-  read_csv(
-    "outputs/data/analyzed_case_data.csv",
-    show_col_types = FALSE)
 ```
@@
 
-```{r}
+```{python}
 #| label: tbl-idcasetable
@@
 
-# Display the first six rows of the "cleaned_case_data" dataset
-head(cleaned_case_data) |>
-  kable(
-    # Give each column a name
-    col.names = c("Case ID", "Age group", "Gender")
+sample = cleaned_case_data.head().rename(
+    columns={
+        "case_id": "Case ID",
+        "age_group": "Age group",
+        "gender": "Gender"
+    }
 )
+
+sample
 ```
@@
 
-```{r}
+```{python}
 #| label: fig-casebygender
-#| fig-cap: Number of covid-19 cases by gender
+#| fig-cap: Number of COVID-19 cases by gender
 #| echo: false
@@
 
-# Create a bar chart to visualize the distribution of COVID-19 cases by gender based on the analyzed_case_data dataset
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
-analyzed_case_data |> 
-  ggplot(aes(x = gender)) +
-  # Change bar color to green
-  geom_bar(fill = "#D4E7C5") +
-  geom_text(
-    # Use the 'count' statistic to calculate counts for each category
-    stat = 'count',
-    # Use percentage labels based on count and set y position of text labels
-    aes(label = scales::percent(..count.. / sum(..count..)),
-        y = ..count..),
-    # Adjust vertical position of text labels
-    vjust = -0.5,
-    # Stack text labels on top of bars
-    position = position_stack(vjust = 0.5)
-  ) +
-  # Apply a minimal theme to the chart
-  theme_minimal() +
-  # Label x-axis as "Gender" and y-axis as "Number of COVID-19 cases"
-  labs(x = "Gender",
-       y = "Number of COVID-19 cases")
+import matplotlib.pyplot as plt
+
+# Count occurrences by gender
+gender_counts = analyzed_case_data["gender"].value_counts()
+gender_percent = gender_counts / gender_counts.sum()
+
+# Plot
+fig, ax = plt.subplots()
+
+bars = ax.bar(gender_counts.index, gender_counts.values, color="#D4E7C5")
+
+# Add percentage labels above bars
+for i, (count, pct) in enumerate(zip(gender_counts.values, gender_percent)):
+    ax.text(i, count + 1000, f"{pct:.1%}", ha='center', va='bottom')
+
+# Label axes and apply minimal formatting
+ax.set_xlabel("Gender")
+ax.set_ylabel("Number of COVID-19 cases")
+ax.set_title("")  # Leave empty if you don't want a title
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+ax.yaxis.grid(True, linestyle='--', alpha=0.4)
+ax.set_axisbelow(True)
+
+plt.tight_layout()
+plt.show()
+
 ```
@@
 
-```{r}
+```{python}
 #| label: fig-casebyagegroup
-#| fig-cap: Number of covid-19 cases by age group
+#| fig-cap: Number of COVID-19 cases by age group
 #| echo: false
@@
 
-# Create a bar chart to visualize the distribution of COVID-19 cases by age group based on the analyzed_case_data dataset
-# Reference: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#explore 
-analyzed_case_data |> 
-  ggplot(aes(x = age_group)) +
-  # Change bar color to green
-  geom_bar(fill = "#D4E7C5") +
-  geom_text(
-    # Use the 'count' statistic to calculate counts for each category
-    stat = 'count',
-    # Use percentage labels based on count and set y position of text labels
-    aes(label = scales::percent(..count.. / sum(..count..)),
-        y = ..count..),
-    # Adjust vertical position of text labels
-    vjust = -0.5,
-    # Stack text labels on top of bars
-    position = position_stack(vjust = 0.5),
-    # Change text size to 3
-    size = 3
-  ) +
-  # Apply a minimal theme to the chart
-  theme_minimal() +
-  # Label x-axis as "Age group" and y-axis as "Number of COVID-19 cases"
-  labs(x = "Age group",
-       y = "Number of COVID-19 cases")
+import matplotlib.pyplot as plt
+
+# Count occurrences by age group and preserve order
+age_order = ["0-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"]
+age_counts = analyzed_case_data["age_group"].value_counts().reindex(age_order)
+age_percent = age_counts / age_counts.sum()
+
+# Plot
+fig, ax = plt.subplots(figsize=(10, 5))
+
+bars = ax.bar(age_counts.index, age_counts.values, color="#D4E7C5")
+
+# Add percentage labels
+for i, (count, pct) in enumerate(zip(age_counts.values, age_percent)):
+    ax.text(i, count + max(age_counts.values) * 0.01, f"{pct:.1%}", ha='center', va='bottom', fontsize=8)
+
+# Label axes and apply minimal styling
+ax.set_xlabel("Age group")
+ax.set_ylabel("Number of COVID-19 cases")
+ax.spines['top'].set_visible(False)
+ax.spines['right'].set_visible(False)
+ax.yaxis.grid(True, linestyle='--', alpha=0.4)
+ax.set_axisbelow(True)
+
+plt.tight_layout()
+plt.show()
+
 ```
@@
 
-```{r}
+```{python}
 #| label: fig-distribution
-#| fig-cap: Distribution of covid-19 cases across genders and age groups
+#| fig-cap: Distribution of COVID-19 cases across genders and age groups
 #| echo: false
@@
 
-# Reference: https://rstudio-pubs-static.s3.amazonaws.com/215516_c2cee6b2885d419e9248e7eb7c87f887.html
-analyzed_case_data |> 
-  # Create a ggplot object with age_group on the x-axis and gender as fill color
-  ggplot(aes(x = age_group, fill = gender)) +
-  # Add a grouped bar layer to create the grouped bar chart
-  geom_bar(position = "dodge") +
-  # Scale the x-axis labels to display "Age group"
-  scale_x_discrete("Age group") +
-  # Scale the fill color to display "Gender" and specify the break points: "Male", "Female", "Other"
-  scale_fill_discrete(name = "Gender",
-                      breaks = c("Male", "Female", "Other")) +
-  # Apply a minimal theme to the chart
-  theme_minimal() +
-  # Label y-axis as "Number of COVID-19 cases"
-  labs(y = "Number of COVID-19 cases")
+import matplotlib.pyplot as plt
+import numpy as np
+
+# --- Prep counts -----------------------------------------------------------
+age_order    = ["0-19", "20-29", "30-39", "40-49", "50-59",
+                "60-69", "70-79", "80-89", "90+"]
+gender_order = ["Male", "Female", "Other"]
+
+counts = (
+    analyzed_case_data
+      .groupby(["age_group", "gender"])
+      .size()                               # tally
+      .unstack(fill_value=0)                # pivot to wide
+      .reindex(age_order)                   # keep age order
+)[gender_order]                              # keep gender order
+
+# --- Plot ------------------------------------------------------------------
+x      = np.arange(len(age_order))
+width  = 0.25                               # width of each bar
+fig, ax = plt.subplots(figsize=(10, 5))
+
+for i, gender in enumerate(gender_order):
+    ax.bar(x + (i - 1)*width,               # dodge position
+           counts[gender],
+           width=width,
+           label=gender)
+
+# Axis / style tweaks
+ax.set_xlabel("Age group")
+ax.set_ylabel("Number of COVID-19 cases")
+ax.set_xticks(x)
+ax.set_xticklabels(age_order)
+ax.legend(title="Gender")
+ax.spines["top"].set_visible(False)
+ax.spines["right"].set_visible(False)
+ax.yaxis.grid(True, linestyle="--", alpha=0.4)
+ax.set_axisbelow(True)
+
+plt.tight_layout()
+plt.show()
+
 ```

### 2025-06-17 outputs/paper/paper.qmd
updated citations

@@
 The data for this paper was obtained from the City of Toronto's Open Data Portal through the 'opendatatoronto' library [@openDataToronto].
-The data analysis involved using the R programming language [@citeR] along with the 'tidyverse' [@citeTidyverse] 'here' [@citeHere], 'janitor' [@citeJanitor], 'knitr' [@citeKnitr] packages for data collection, cleaning, and analysis.
-Within the 'tidyverse' package [@citeTidyverse], I used 'readr' [@citeReadr], 'dplyr' [@citeDplyr], 'tidyr' [@citeTidyr], and 'ggplot2' [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating bar charts.
+The data analysis involved using the R programming language [@python] along with the 'numpy' [@numpy] and 'pandas' [@pandas] packages for data collection, cleaning, and analysis.
+'matplotlib' [@matplotlib] was used for creating bar charts.
 

### 2025-06-23 outputs/paper/paper.qmd
quick fix
anonymized & changed packages

@@
 ---
-title: "Exploring the Demographic Aspects of COVID-19 Cases in Toronto: Are the elderly males more susceptible to COVID-19?"
+title: "Demographic Patterns in Toronto’s COVID-19 Cases: Are Elderly Males at Greater Risk?"
 author: 
-  - Amie Liu
-thanks: "Code and data are available at:https://github.com/AmieLiu/COVID-19_Toronto_Demographic_Aspects_Analysis.git"
+  - Lena Wang
+thanks: "Code and data are available at: https://github.com/EveHughes/COVID-19_Toronto_Demographic_Aspects_Analysis"
 date: today
@@
 
-The dataset 'COVID-19 Cases in Toronto'  [@covidCases] was utilized and obtained from 'opendatatoronto' library [@openDataToronto].
+The dataset 'COVID-19 Cases in Toronto' [@covidCases] was utilized.
 It comprises 412,122 records and 15 variables, including assigned ID, age group, neighborhood name, episode date, client gender, and more.

### 2025-06-23 scripts/00-simulate_data.py

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Simulates the dataset of COVID-19 cases in Toronto
-# Author: Amie Liu
+# Author: Lena Wang
 # Date: 16 January 2024
-# Contact: amie.liu@mail.utoronto.ca
+# Contact: lena.wang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: None
 

### 2025-06-23 scripts/01-download_data.py

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from the Open Data Toronto portal
-# Author: Amie Liu
+# Author: Lena Wang
 # Date: 16 January 2024
 # Contact: amie.liu@mail.utoronto.ca
 # License: MIT
@@ -9,19 +9,12 @@
 #### Workspace setup ####
 import pandas as pd
 from io import StringIO
-from datetime import datetime
 
 #### Download COVID-19 cases in Toronto data ####
 import requests
 
-# Toronto Open Data is stored in a CKAN instance. It's APIs are documented here:
-# https://docs.ckan.org/en/latest/api/
-
-# To hit our API, you'll be making requests to:
 base_url = "https://ckan0.cf.opendata.inter.prod-toronto.ca"
 
-# Datasets are called "packages". Each package can contain many "resources"
-# To retrieve the metadata for this package and its resources, use the package name in this page's URL:
 url = base_url + "/api/3/action/package_show"
 params = { "id": "covid-19-cases-in-toronto"}
 package = requests.get(url, params = params).json()

### 2025-06-23 scripts/02-data_cleaning.py

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Cleans the raw COVID-19 cases data
-# Author: Amie Liu
+# Author: Lena Wang
 # Date: 16 January 2024
-# Contact: amie.liu@mail.utoronto.ca
+# Contact: lena.wang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites:
 #   00-simulate_data.py
@@ -10,7 +10,6 @@
 
 #### Workspace setup ####
 import pandas as pd
-import numpy as np
 
 #### Clean data ####
 # Read raw data

### 2025-06-23 scripts/03-test_data.py

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Test cleaned COVID-19 cases data
-# Author: Amie Liu
+# Author: Lena Wang
 # Date: 25 January 2024
-# Contact: amie.liu@mail.utoronto.ca
+# Contact: lena.wang@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
 #   00-simulate_data.py
